Validation of an Actuator Line Model Coupled to a Dynamic Stall Model for Pitching Motion Characteristic to Vertical Axis Turbine

### 2016-02-29 paper.tex
Minor formatting edits

\section{Methodology}
For solving the governing equations of phenomena involved, a coupled model have been implemented: the actuator line model is used to calculate the flow velocities and thereby the angle of attack. The dynamic stall model is used to calculate blade force coefficients, which the actuator line model need for the velocity field calculations. In this work it was chosen to validate the model against wind tunnel data for a pitching blade. This will put the focus on the force modeling part in the simulation model\\
\\
\textbf{Actuator Line Model:} is a three-dimensional and unsteady aerodynamic model developed by S{\o}rensen and Shen\cite{sorensen1999computation}, used to study the flow around wind turbines. It is a combination of a solver of Navier-Stokes equations with a so-called actuator line technique, in which blades of the turbine are represented by a radial distribution of body forces along lines. These forces are determined using a dynamic stall model commonly based on empirical data.\\
\\
\textbf{Dynamic Stall Modeling:} represented by the Leishman-Beddoes model\cite{leishman1986generalised} with the modifications of Sheng et al\cite{sheng2008modified}. It is capable to calculate the unsteady lift, pitching moment and drag, giving a physical description of the aerodynamics. It have been widely validated with experimental data\cite{leishman1989semi}. Separated on three subsystems: an attached flow model for unsteady linear airloads, a separated flow model for non-linear airloads and a dynamic stall model for the airloads induced by the leading edge vortex. In this work, the dynamic stall model described by Dyachuk\cite{dyachuk} has been implemented, using the library turbinesFoam developed by Bachant\cite{bachant2015simulating}.\\
\\
NACA0021 airfoil was tested at the Reynolds number of 1000000 (which is a reasonable value for operating VAWT) during pithing motion similar to the motion of the VAWT blade. Different amplitude degrees were tried out. In a VAWT with fixed blades the angle of attack is in function of the TSR of the turbine. A periodic function is used to represent the variations of the angles of attack which a VAWT blade experiences analog to the pitching motion,
\begin{eqnarray}
\alpha = \arctan \left( \frac{\sin \theta}{\lambda + \cos \theta} \right)
\end{eqnarray}
Where $\theta$ is the azimuthal blade angle and $ \lambda $ represents the TSR,
\begin{eqnarray}
\lambda =  \frac{\Omega R}{V}
\end{eqnarray}
The pitching blade experiments were carried out at Glasgow University\cite{angell1988collected}\\

\section{Results and discussion}
The reliability of the model results is acceptable, making a good agreement with experimental results considering the complexity of the flows studied. For a TSR of 3.34 (Figure \ref{fig1}), the maximum magnitude of the angle of attack is at 17.4\degree which is in the stall region. Both CN and CT curves show the delay of the flow reattachment, which is characteristic of the dynamic stall phenomenon. Peaks of simulated values have close agreement with the measured data.\\
For a TSR of 2.60, the maximum amplitude of the angle of attack is 22.6, which is related to a deeper stall region compared to a TSR of 3.34. This is shown because the "loop" of the curve is wider for the lower TSR, and moreover, the reattachment of the flow is further delayed.\\
It have been demonstrated how the coupled actuator line model can be coupled to a modified version of the Leishman-Beddoes model, and the results could be validated for similar operating conditions of a VAWT.

>>>>

\section{Methodology}

For solving the governing equations of phenomena involved, a coupled model have
been implemented: the actuator line model is used to calculate the flow
velocities and thereby the angle of attack. The dynamic stall model is used to
calculate blade force coefficients, which the actuator line model need for the
velocity field calculations. In this work it was chosen to validate the model
against wind tunnel data for a pitching blade. This will put the focus on the
force modeling part in the simulation model.

\textbf{Actuator Line Model:} is a three-dimensional and unsteady aerodynamic
model developed by S{\o}rensen and Shen\cite{sorensen1999computation}, used to
study the flow around wind turbines. It is a combination of a solver of
Navier-Stokes equations with a so-called actuator line technique, in which
blades of the turbine are represented by a radial distribution of body forces
along lines. These forces are determined using a dynamic stall model commonly
based on empirical data.

\textbf{Dynamic Stall Modeling:} represented by the Leishman-Beddoes
model\cite{leishman1986generalised} with the modifications of Sheng et
al\cite{sheng2008modified}. It is capable to calculate the unsteady lift,
pitching moment and drag, giving a physical description of the aerodynamics. It
have been widely validated with experimental data\cite{leishman1989semi}.
Separated on three subsystems: an attached flow model for unsteady linear
airloads, a separated flow model for non-linear airloads and a dynamic stall
model for the airloads induced by the leading edge vortex. In this work, the
dynamic stall model described by Dyachuk\cite{dyachuk} has been implemented,
using the library turbinesFoam developed by Bachant\cite{bachant2015simulating}.

NACA0021 airfoil was tested at the Reynolds number of 1000000 (which is a
reasonable value for operating VAWT) during pithing motion similar to the motion
of the VAWT blade. Different amplitude degrees were tried out. In a VAWT with
fixed blades the angle of attack is in function of the TSR of the turbine. A
periodic function is used to represent the variations of the angles of attack
which a VAWT blade experiences analog to the pitching motion,
\begin{eqnarray}
    \alpha = \arctan \left( \frac{\sin \theta}{\lambda + \cos \theta} \right)
\end{eqnarray}
Where $\theta$ is the azimuthal blade angle and $ \lambda $ represents the TSR,
\begin{eqnarray}
    \lambda =  \frac{\Omega R}{V}
\end{eqnarray}
The pitching blade experiments were carried out at Glasgow
University\cite{angell1988collected}\\

\section{Results and discussion}

The reliability of the model results is acceptable, making a good agreement with
experimental results considering the complexity of the flows studied. For a TSR
of 3.34 (Figure \ref{fig1}), the maximum magnitude of the angle of attack is at
17.4\degree which is in the stall region. Both CN and CT curves show the delay
of the flow reattachment, which is characteristic of the dynamic stall
phenomenon. Peaks of simulated values have close agreement with the measured
data.

For a TSR of 2.60, the maximum amplitude of the angle of attack is 22.6, which
is related to a deeper stall region compared to a TSR of 3.34. This is shown
because the ``loop'' of the curve is wider for the lower TSR, and moreover, the
reattachment of the flow is further delayed.

It has been demonstrated how the coupled actuator line model can be coupled to
a modified version of the Leishman-Beddoes model, and the results could be
validated for similar operating conditions of a VAWT.
